Unsupervised embeddings of organizations capture latent structure of scientific mobility

### 2020-03-31 paper/current/main.tex
revise + re-organize data and methods

Data, analysis, and associated code have been made publicly available and can be found at \textit{https://github.com/murrayds/...}

\subsection*{Data}
%% Main dataset
% Do we need to add the Leiden field classificaiton infomration? It doesn't seem like its relevant for the current analysis.
% Do we need to talk about how we define mobility in relation to past papers?
We source co-affiliation trajectories of authors from the Web of Science database hosted by the Center for Science and Technology Studies at Leiden University.
Trajectories are constructed from author affiliations listed on the byline of publications for a given author.
Given the limitations of author-name disambiguation, we limit to papers published after 2008, when the Web of Science began providing full names and institutional affiliations~\autocite{caron2014disambiguation} and when performance is strongest (ref to SI?).
This yields 33,934,672 author-affiliation combinations representing 12,963,792 authors.
Each author-affiliation combination is associated with the publication year and a unique id that maps the affiliation to a disambiguate organization.
Combinations are also associated with a publication-level disciplinary classification, one of ``Social sciences and humanities'',
``Biomedical and health sciences'', ``Physical sciences and engineering'', ``Life and earth sciences'', and ``Mathematics and computer science''.
Trajectories are represented as the list of author-affiliation combinations, ordered by year of publication, and randomly ordered for combinations within the same year.
Authors are classified as mobile when they have at least two distinct organization ids in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.
While the number of publications has increased year-to-year, the mobility and disciplinary makeup of the dataset has not significantly changed across the period of study (Fig.~\ref{fig:supp:pubs_over_time}).

% %Organization information
Affiliations mapped to one of 8,661 organizations, disambiguated following a mostly manual process originally designed for the Leiden Rankings of World Universities~\autocite{waltman2012leidenrankings}.
Organizational records were associated with a full name, a type indicating the sector (e.g., University, Government, Industry), and an identifier for the country and city of the organization.
Sixteen different sector types were included in the analysis, which we aggregated to a four high-level codes: \textit{University}, \textit{Hospital}, \textit{Government}, and \textit{Other}.
Each record was also associated with a latitude and longitude, however for many organizations these were missing or incorrect. We manually updated the coordinates of 2,267 organizations by searching the institution name and city on Google Maps;
in cases where a precise location of the organization could not be identified, we used the coordinates returned when searching the name of the city.
The data was further enriched with country-level information, including continent, majority language spoken, and language family of the majority language (e.g., the language family of \textit{Spanish} is \textit{Italic}).
State/province level  information were added using the reverse geocoding service LocationIQ using each organization's latitude and longitude as input.
Regional census classifications were also added for states in the United States.
For each organization, we also calculated a size as the average number of unique authors (mobile and non-mobile) who published with that organization across each year of our dataset;
in the case that authors publish with multiple affiliations in a single year, they are counted towards each.

Scientific mobility is complex and can be defined in many possible ways~\autocite{robinson2019mobility}.
Here, we consider an author as mobile when at least two distinct organizational affiliations appear in their trajectory during the 2008-2019 time window.
Under this definition, mobile authors constitute 3,007,192 or 23.2 percent of all authors, and 17,700,095 author-affiliation combinations.
Mobile authors were associated with, on average, 2.5 distinct organizational affiliations.
Rates of mobility tended to differ across country.
For example, France, Qatar, the USA, Iraq, and Luxembourg had the most mobile authors (ref figure).
However, due to their size, the USA, accounted for nearly 40 percent of all mobile authors worldwide, with  10 countries accounting for 80 percent of all mobility (ref figure).
In most cases, countries with a high degree inter-organization mobility also had a high degree of international mobility;
however, the USA and France had a high degree of inter-organization mobility with the corresponding international mobility (ref figure).

@@

%
% SUBSECTION: DISTANCE
%
\subsection{Distance}
%% Distance calculations
We calculate $T_{ij}$, the pairwise number of co-affiliations between organizations, calculated as the number of co-occurrences of organizations within author co-affiliation trajectories.
For example, $T_{ij} = 10$ indicates that ten authors were co-affiliated with both organization $i$ and $j$ between 2008 and 2019, as evidenced from their publications.
In contrast to traditional mobility studies, all co-affiliations are symmetric, meaning that $T_{ij} = T_{ji}$.

@@

%
% SUBSECTION: GRAVITY LAW
%
\subsection{Gravity Law}

@@

\begin{equation}
	\label{eq:ppr}
	p_i = (1.0 - \alpha) * v_i + \alpha p_i W,
\end{equation}

@@

@@

%
% Figure - publications over time
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Descriptive/mobility_by_country.pdf}
	\caption{
		\textbf{Extent and nature of mobility by country.}
		\textbf{a.}
		Proportion of researchers in each country classified as mobile, meaning that they have at least two organization affiliations within the period of study.
		The countries with the most mobile researchers are the United States, China, France, Great Britain, and Germany.
		\textbf{b.}
		The same information, arranged into a cumulative distribution.
		The five countires with the highest mobility account for approximately 70 percent of all mobility.
		\textbf{c.} The proportion of researchers labeled as mobile by country.
		\textbf{d.} First two principal components of four variables: proportion in each country mobile across organizations, proportion mobile across cities, proportion mobile across regions, and proportion mobile across countries.
		First component captures overall degree of mobile, whereas the second compares within-country against across-country mobility.
	}
	\label{fig:supp:pubs_over_time}
\end{figure}

>>>>

Data, analysis, and associated code have been made publicly available and can be found at \url{https://github.com/murrayds/sci-mobility-emb}

\subsection*{Data}
%% Main dataset
% Do we need to talk about how we define mobility in relation to past papers?
We source co-affiliation trajectories of authors from the Web of Science database hosted by the Center for Science and Technology Studies at Leiden University.
Trajectories are constructed from author affiliations listed on the byline of publications for a given author.
Given the limitations of author-name disambiguation, we limit to papers published after 2008, when the Web of Science began providing full names and institutional affiliations~\autocite{caron2014disambiguation} that improved disambigatuon (see supporting materials).
This yields 33,934,672 author-affiliation combinations representing 12,963,792 authors.
Each author-affiliation combination is associated with the publication year and an ID linking it to one of 8,661 a dismabiguated organizational affiliation (see supporting materials for more detail)
Trajectories are represented as the list of author-affiliation combinations, ordered by year of publication, and randomly ordered for combinations within the same year.

Scientific mobility is complex and can be defined in many possible ways~\autocite{robinson2019mobility}.
Here, authors are classified as mobile when they have at least two distinct organization ids in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.
Under this definition, mobile authors constitute 3,007,192 or 23.2 percent of all authors, and 17,700,095 author-affiliation combinations.
Mobile authors were associated with, on average, 2.5 distinct organizational affiliations.
Rates of mobility tended to differ across country.
For example, France, Qatar, the USA, Iraq, and Luxembourg had the most mobile authors (Fig.~\ref{fig:supp:descriptives}c).
However, due to their size, the USA, accounted for nearly 40 percent of all mobile authors worldwide (Fig.~\ref{fig:supp:descriptives}a), with 10 countries accounting for 80 percent of all mobility (Fig.~\ref{fig:supp:descriptives}b).
In most cases, countries with a high degree inter-organization mobility also had a high degree of international mobility;
however, the USA and France had a high degree of inter-organization mobility with the corresponding international mobility (Fig.~\ref{fig:supp:descriptives}c).
While the number of publications has increased year-to-year, the mobility and disciplinary makeup of the dataset has not significantly changed across the period of study (Fig.~\ref{fig:supp:pubs_over_time}).

@@

%
% SUBSECTION: DISTANCE
%
\subsection*{Distance}
%% Distance calculations
We calculate $T_{ij}$, the pairwise number of co-affiliations between organizations, calculated as the number of co-occurrences of organizations within author co-affiliation trajectories.
For example, $T_{ij} = 10$ indicates that ten authors were co-affiliated with both organization $i$ and $j$ between 2008 and 2019, as evidenced from their publications.
In contrast to traditional mobility studies, all co-affiliations are symmetric, meaning that $T_{ij} = T_{ji}$.

@@

%
% SUBSECTION: GRAVITY LAW
%
\subsection*{Gravity Law}

@@

\begin{equation}
	\label{eq:ppr}
	p_i = (1.0 - \alpha) v_i + \alpha p_i W,
\end{equation}

@@

%
% S2 TEXT
\paragraph*{S2 Text}
\label{si:text:organizations}
{\bf Organization disambiguation and metadata.}

Affiliations mapped to one of 8,661 organizations, disambiguated following that origionally designed for the Leiden Rankings of World Universities~\autocite{waltman2012leidenrankings}.
Organizational records were associated with a full name, a type indicating the sector (e.g., University, Government, Industry), and an identifier for the country and city of the organization.
Sixteen different sector types were included in the analysis, which we aggregated to a four high-level codes: \textit{University}, \textit{Hospital}, \textit{Government}, and \textit{Other}.
Each record was also associated with a latitude and longitude, however for many organizations these were missing or incorrect. We manually updated the coordinates of 2,267 organizations by searching the institution name and city on Google Maps;
in cases where a precise location of the organization could not be identified, we used the coordinates returned when searching the name of the city.
The data was further enriched with country-level information, including continent, majority language spoken, and language family of the majority language (e.g., the language family of \textit{Spanish} is \textit{Italic}).
State/province level  information were added using the reverse geocoding service LocationIQ using each organization's latitude and longitude as input.
Regional census classifications were also added for states in the United States.
For each organization, we also calculated a size as the average number of unique authors (mobile and non-mobile) who published with that organization across each year of our dataset;
in the case that authors publish with multiple affiliations in a single year, they are counted towards each.

%
% S3 TEXT
\paragraph*{S3 Text}
\label{si:text:disambiguation}
{\bf Author name dismabiguation.}

Author-name disambiguation, the problem of associating names on papers with individuals authors, remains difficult for the use of bibliometric data~\autocite{dangelo2020disambiguation}. 
Authors in our dataset have been disambiguated using a rule-based algortihm that makes use of author and paper metadata, such as physica, addresses, co-authors, and journal, to score papers on the likelihood of belonging to an author cluster—a cluster of publications belived to have been authored by the same individual~\autocite{caron2014disambiguation}. 
We limit our period of analysis until after 2008, as this is when the Web of Science began indexing additional author-level metadata such as full names and email addresses. 
The disambiguation algorithm is conservative, favoring splitting clusters. 
Past studies have validated this data, demonstrating it comparable to ground-truth records such as those from ORCID and useful for a wide range of bibliometric studies~\cite{sugimoto2017mostimpact, robinson2019mobility, chinchilla2018global, chinchilla2018travelban}.

\newpage

@@

%
% Figure - descriptives
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Descriptive/mobility_by_country.pdf}
	\caption{
		\textbf{Extent and nature of mobility by country.}
		\textbf{a.}
		Proportion of researchers in each country classified as mobile, meaning that they have at least two organization affiliations within the period of study.
		The countries with the most mobile researchers are the United States, China, France, Great Britain, and Germany.
		\textbf{b.}
		The same information, arranged into a cumulative distribution.
		The five countires with the highest mobility account for approximately 70 percent of all mobility.
		\textbf{c.} The proportion of researchers labeled as mobile by country.
		\textbf{d.} First two principal components of four variables: proportion in each country mobile across organizations, proportion mobile across cities, proportion mobile across regions, and proportion mobile across countries.
		First component captures overall degree of mobile, whereas the second compares within-country against across-country mobility.
	}
	\label{fig:supp:descriptives}
\end{figure}
